Beating Simple Backtracking Search in Subgraph Algorithms\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1].}

### 2018-11-27 paper/paper.tex
Are we nearly there yet?

Although the problem is \NP-complete, by combining design techniques from artificial intelligence
with careful algorithm engineering, modern subgraph isomorphism solvers can often produce exact
solutions quickly even on graphs with thousands of vertices. The current single strongest subgraph
isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to direct a
constraint programming style search
\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:journals/jair/McCreeshPST18}. This
heuristic is much better than branching randomly, but is still far from perfect. To offset mistakes
made by this heuristic, this paper proposes a new perspective on value-ordering: rather than
defining a search order, we use degree to direct what \emph{proportion} of the search effort should
be spent in each subproblem. By combining rapid restarts and nogood recording, and introducing a
small amount of randomness into the value-ordering heuristic, we make a state-of-the-art subgraph
algorithm perform two orders of magnitude better on a large number of satisfiable instances, whilst
performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.  This
strategy is also effective in an optimisation setting, producing benefits in two maximum common
induced subgraph algorithms. Through this combination of techniques we have, for the first time, a
practical alternative to backtracking search for constraint programming style algorithms where we
have a strong value-ordering heuristic and where we care both about satisfiable and unsatisfiable
instances.

@@

Although we have shown that it provides good results, we have yet to justify where the biased
formula comes from, or indeed why we call this approach ``solution-biased''.  Our goal is to use
biased randomness in a value-ordering heuristic to spend time in subproblems proportional to an
estimate of their solution density \cite{DBLP:journals/jair/PesantQZ12}. ?? I feel like we still
need a sentence here saying why this is a good idea.

@@

We have shown that it is both possible and beneficial to introduce a small amount of randomness into
backtracking search algorithms for hard subgraph problems, without compromising existing search
order heuristics, and without paying a penalty for unsatisfiable instances. In principle it is
well-known that introducing some form of randomness \cite{DBLP:conf/aaai/GomesSK98} or reduced
commitment \cite{DBLP:conf/ijcai/HarveyG95} into backtracking search can be beneficial, but until
now it has not been clear how to do so efficiently in practice, particularly when we already have
strong search ordering heuristics and when we are expecting to encounter large numbers of unsatisfiable
instances. Our results show that the key to success is a combination of aggressive restarts, nogood
recording, and biasing an existing value-ordering heuristic with just a pinch of randomness.

?? Parallel

?? Apply to CP in general

% \section*{Acknowledgements}
% 
% The authors would like to thank Ozgur Akgun, Chris Jefferson, Christophe Lecoutre, Christine
% Solnon, and Craig Reilly for their comments.

>>>>

Although the problem is \NP-complete, by combining design techniques from artificial intelligence
with careful algorithm engineering, modern subgraph isomorphism solvers can often produce exact
solutions quickly even on graphs with thousands of vertices. The current single strongest subgraph
isomorphism solver uses ``highest degree first'' as a value-ordering heuristic to direct a
constraint programming style search
\cite{DBLP:conf/cp/McCreeshP15,DBLP:conf/lion/KotthoffMS16,DBLP:journals/jair/McCreeshPST18}. This
heuristic is much better than branching randomly, but is still far from perfect. To offset mistakes
made by this heuristic, this paper proposes a new perspective on value-ordering: rather than
defining a search order, we use degree to direct what \emph{proportion} of the search effort should
be spent in each subproblem. By combining rapid restarts and nogood recording, and introducing a
small amount of randomness into the value-ordering heuristic, we make a state-of-the-art subgraph
algorithm perform two orders of magnitude better on a large number of satisfiable instances, whilst
performing worse only rarely on satisfiable instances, and never on unsatisfiable instances.  This
strategy is also effective in an optimisation setting, producing benefits in two maximum common
induced subgraph algorithms.

@@

Although we have shown that it provides good results, we have yet to justify where the biased
formula comes from, or indeed why we call this approach ``solution-biased''.  Our goal is to use
biased randomness in a value-ordering heuristic to spend time in subproblems proportional to an
estimate of their solution density \cite{DBLP:journals/jair/PesantQZ12}. Such an approach is better
than committing entirely to the area of maximum solution density because estimators only give a
probability---although we may estimate that one subtree has twice the solution density of another,
in reality the ``better'' subtree may contain no solutions at all.

@@

The conventional view of value-ordering heuristics is that they define a search order. We have
proposed an alternative perspective, where value-orderings define a weighting specifying how much
search effort should be put into different subproblems, based upon a rough estimate of solution
densities. We have also shown how to turn this perspective into an algorithm, by combining a biased
random value-ordering heuristic with rapid restarts and nogood recording. This combination of
techniques gives us, for the first time, a practical alternative to backtracking search where we
have a strong value-ordering heuristic and where we care both about satisfiable and unsatisfiable
instances.  A further benefit is the ease with which such a search can be parallelised. By having
each thread carry out the same search with a different random seed, and sharing nogoods only on
restarts, we remove the need for intrusive changes to the core search algorithm, eliminate the
irregularity problem, and still respect the advice of the value-ordering heuristic.

We believe that these technique are broadly applicable, beyond subgraph algorithms, and we intend to
implement them in a full constraint programming solver. We are also interested in making better use
of statistical knowledge (either \emph{a priori} or learned during search) to further refine the
biased randomisation process. And finally, we are trying hard to work out whether our new
perspective also has some relevance to \emph{variable} ordering heuristics.

\section*{Acknowledgements}

The authors would like to thank Ozgur Akgun, Chris Jefferson, Christophe Lecoutre, Christine
Solnon, and Craig Reilly for their comments.
